---
title: "a3_task3_sachiko_lamen"
author: "Sachiko Lamen"
date: "10/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, messages = FALSE, warnings = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(kableExtra)
```

This document wrangles and visualizes data for red and purple urchins recorded by the Santa Barbara Long Term Ecological Research (LTER).

**Citation:** Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. <https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b>.

Variable Descriptions:

-   date: date urchin recorded

-   site: site along the SB coast (the five sites can be explored here)

-   treatment: indicates experimental kelp treatments:

    -   CONTROL = giant kelp left intact
    -   ANNUAL = giant kelp is "removed...once each winter and allowed to re-colonize over the course of the year"

-   size: urchin diameter (cm), measured to the nearest 0.5 cm

-   scientific_name: scientific name

-   common_name: species common name (either red or purple urchin)

### Read in/clean data

First 6 lines are shown for reference

```{r}
urchins <- read_csv(here("data", "urchins.csv")) %>% clean_names()

head(urchins)
```

### Size Comparison: Data Visualization

Comparing the distribution of sizes (cm) between purple and red urchins at all sites from 2007 - 2020.

```{r}

ggplot(data = urchins, aes(x = size, fill = common_name)) +
    geom_histogram( color="#e9ecef", alpha=0.4, position = 'identity', bins = "10") +
    scale_fill_manual(values = c("#404080", "red")) +
  facet_wrap(~ common_name) +
  labs(x = "Urchin Size (cm)",
       y = "Number of Urchins",
       title = "Urchin Size Comparison") + 
  theme(legend.position = "none")
```

**Figure 1. Total size distribution of purple and red urchins in Santa Barbara Channel (2008-2020).** Urchin size (cm) data was collected from five sites: AQUE (n = 12480), CARP (n = 13772), IVEE (n = 6771), MOHK (n = 12306), NAPL (n = 13884). Bars indicate total number of urchins recorded at each size (bins = 10).

### Size Comparison: Descriptive Statistics Table

```{r}
urchin_stats <- urchins %>%
  group_by(common_name) %>%
  summarize(mean_size = mean(size, na.rm = TRUE),
           median_size = median(size, na.rm = TRUE),
            std_dev = sd(size, na.rm = TRUE),
           sample_size = n())


urchin_stats %>%
  kable(col.names = c("Common Name",
                      "Mean Body Size (cm)", 
                      "Median Body Size (cm)",
                      "Standard Deviation",
                      "Sample Size (n)"),
        caption = "**Table 1. Species Specific Size Statistics.** Data was collected from five sites: AQUE (n = 12480), CARP (n = 13772), IVEE (n = 6771), MOHK (n = 12306), NAPL (n = 13884) by Santa Barbara Long Term Ecological Research divers (2008 - 2020). Urchin size was measured using diameter (cm).") %>% 
  kable_styling(full_width = FALSE)
```

Averaging across all treatments and sites, the average diameter of red urchins (6.78 cm) is larger 34% larger than the average diameter of purple urchins (4.47 cm). This difference is supported by the large difference in median values; red urchin (7.0 cm), purple urchin (4.5 cm). The red urchin data is more widely distributed (sd = 1.89) than the purple urchin data (sd = 1.18).

### Purple Urchin Counts

Create stored subset `purple_urchins` that shows only data for purple urchins. Convert date to class "Date" and add new columns `year` and `site_full` that contains full name of site based on acronym.

```{r}
purple_urchins <- urchins %>%
  filter(common_name == "Purple Urchin") %>%
  mutate(date = mdy(date)) %>%
  mutate(year = year(date)) %>%
  mutate(site_full = case_when(
    site %in% c("NAPL") ~ "Naples Reef",
    site %in% c("AQUE") ~ "Arroyo Quemado Reef",
    site %in% c("CARP") ~ "Carpinteria Reef",
    site %in% c("MOHK") ~ "Mohawk Reef",
    site %in% c("IVEE") ~ "Isla Vista Reef"))

```

Create subset `total_number_purple_urchins` with urchin counts grouped by site and year.

```{r}
total_number_purple_urchins <- purple_urchins %>%
  group_by(site, year) %>%
  count(common_name)

```


```{r}
ggplot(data = total_number_purple_urchins, aes(x = year, y = n)) +
  geom_point(aes(color = site),
             size = 0.6) +
  scale_x_continuous(breaks = seq(from = 2007, to = 2020, by = 1)) +
  labs(x = "Year",
      y = "Number of Urchins",
      title = "Purple Urchins Recorded at Each Site")
```
**Figure 2. Yearly purple urchin counts at five sites in Santa Barbara Channel (2008-2020).** Urchin diameter size (cm) data was collected from five sites: Arroyo Quemado Reef (AQUE), Carpinteria Reef (CARP), Isla Vista Reef (IVEE), Mohawk Reef (MOHK), Naples Reef (NAPL). Individual points represent total number of urchins recorded at each site/year across both treatments.

According to this graph, it appears that purple urchin population is in decline. We need to take into account that there are different treatments, `CONTROL`, and `ANNUAL`. We would want to visualize/analyze the data in such a way that control sites are separated/compared to annual sites to get a better idea of population dynamics. Additionally we do not know the rate that urchins are harvested in Santa Barbara. If we knew how many urchins were collected by professional/recreational urchin divers at each site we would have a better understanding of true urchin population. It is possible that as dive/boat technology has improved, harvesting pressure has increased while urchin population has remained constant.
